\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-09-03 archive.tex
final poster
compression notes
edits to introduction

\usepackage{dblfloatfix}    % To enable figures at the bottom of page
% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

\usepackage[section]{placeins}
% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
%\usepackage{hyperref}

@@

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\scode}[1]{{\tiny\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

\section{Introduction}

@@

  \end{figure}
\begin{figure}
\centering
\tiny
\tabcolsep=2pt
\vspace{0pt}
\begin{tabular}{cc}
\begin{tabular}{ccc}
\multicolumn{3}{c}{Tasks:}\\
\hline
cut&F &Moss Side \\ 
control&CL &Burnage \\
control&F & City Centre\\
cut& PCFL& Brooklands \\
\multicolumn{3}{c}{Learned generative models:}\\
\hline
\scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash
u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash
u\textbackslash w*)*} \\
\multicolumn{3}{c}{Samples from synthesized generative models:}\\
\hline
ya&DQDF & Vr DR \\ 
glrwfdcnc&F &BeF lKQ \\
mgs&F& W\\
piljnl& KI& kqBfZ 0 \\
kj& F& ON\\
zci& GL& Bttc \\
sxpm& F& S \\
\hline
\\
\end{tabular} 
&
\begin{tabular}{c}
\begin{tabular}{l}
\multicolumn{1}{c}{Learned DSL:}\\
\hline
$f_1() = $ \scode{\textbackslash u\textbackslash w*}\\
$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash
u\textbackslash w*)*}\\
$f_3(x) = f_2(\scode{space}) = $\scode{( \textbar\textbackslash
u\textbackslash w*)*}\\
\hline
$f_4(x) = $ \scode{($x$*$x$)}\\
\emph{ (equivalent to regex `plus')}\\
$f_5() = f_4(\scode{\textbackslash l}) = \scode{\textbackslash
l*\textbackslash l}$\\
\hline
\end{tabular}
\\
\\
\includegraphics[width=.4\linewidth]{figures/nampiregexjoshcurve3.eps} 
\end{tabular}
\end{tabular}
\vspace{-.3cm}
\caption{regex stuff}
\end{figure}

>>>>

\usepackage{dblfloatfix}    % To enable figures at the bottom of page
% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs} \usepackage{stmaryrd}% for professional tables
\usepackage{stmaryrd}(
\usepackage[section]{placeins}
% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
%\usepackage{hyperref}

@@

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\denotation}[1]{{\llbracket #1 \rrbracket}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\scode}[1]{{\tiny\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

\section{Introduction: Expertise}

Human learners acquire expertise in a wide range of domains: some of
us become experts in calculus, or cooking, or biology, music, tennis,
or software engineering, to name just a few, and every child develops
expertise in natural language, intuitive physics~\cite{Pete?},
intuitive psychology (theory-of-mind), and motor control.  This
picture contrasts sharply with the current state of machine
intelligence, where a machine is built to be an expert in a single
domain, like boardgames~\cite{alphaGo}, medical
diagnosis~\cite{mycin,CNNsForDiseaseDetection}, theorem
proving~\cite{eurisko}, or visual object recognition~\cite{alexNet}.
Thus an outstanding challenge in the long-term program of building
more humanlike machines is to develop an algorithm that, like people,
autonomously acquires expertise across many different kinds of
domains.

We contribute a model of the development of expertise that combines
two key ingredients.  First, an expert needs a sufficiently expressive
knowledge representation.  Following a long tradition in cognitive
science and AI, we represent knowledge as programs, which prior work
has used to represent expertise in recognition and generation of
handwriting and speech~\cite{lake2015human}, intuitive theories (of
kinship, taxonomy, etc.)~\cite{Ullman2012}, and natural language
grammar and
semantics~\cite{DBLP:journals/cogsr/SchmidK11,piantadosi2011learning}.
Second, atop this program representation, experts possess two kinds of
domain expertise.  They have at their disposal a powerful, yet
specialized, repertoire of concepts and abstractions: e.g., in
architecture, these are concepts like `arch' or `foundation'; in
software engineering these are libraries of code and domain-specific
languages (DSLs).  Here our model represents knowledge as programs,
and so we identify this kind of expertise with a DSL.  Experts also
have knowledge of when and how to deploy these domain-specific
concepts efficiently when solving new problems: e.g., mathematicians
efficiently search the space of proofs, intuiting which lemmas are
appropriate when; expert chefs intuit which compositions of
ingredients are likely tasty, before they actually start cooking.  For
our model, this aspect of expertise corresponds to the ability to
quickly assemble new, useful programs out of its DSL.

We integrate these ideas into a model called DreamCoder which acquires
expertise through a novel kind of wake/sleep or `dream' learning. The
model iterates through wake cycles -- where it solves problems by
writing programs -- and a pair of sleep cycles: a sleep cycle that
grows its DSL by replaying experiences from waking and consolidating
them into new abstractions, and a sleep cycle that improves its
knowledge of how to write programs by training a neural network on
replayed experiences as well as `dreams', or samples, from its DSL.

DreamCoder builds on multiple generations of AI research, going back
to the 1960's~\cite{solomonoff1964formal} when program-learning was
proposed as a paradigm for general AI.  Broadly speaking recent work
has either developed neural approaches for learning to efficiently
deploy a fixed
DSL~\cite{devlin2017robustfill,balog2016deepcoder,NGDS,spiral}, or
developed symbolic approaches for representing and searching through
spaces of
programs~\cite{gulwani2011automating,solar2008program,DBLP:books/daglib/0070933}.
We were motivated by approaches that learn or grow the
DSL~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning}.
Our goal with DreamCoder is to show that the combination of
neurally-guided search and DSL learning is a uniquely powerful way of
building systems that, like human learners, autonomously acquire the
expertise needed to navigate a new domain of problems.

\section{Introduction: AI/PL/dreaming}

An age-old dream within AI is a machine that learns and reasons by
writing its own programs.  This vision stretches back to the
1960's~\cite{solomonoff1964formal} and cuts to the core of much of
what it would take to build machines that learn and think like humans.
Computational models of cognition often explain the flexibility and
richness of human thinking in terms of program learning: from everyday
thinking and problem solving (motor program induction as an account of
recognition and generation of handwriting and
speech~\cite{lake2015human}; functional programs as a model of natural
language semantics~\cite{SOMETHING}) to learning problems that unfold
over longer developmental time scales: the child's acquisition of
intuitive theories (of kinship, taxonomy, etc.)~\cite{Ullman2012} and
natural language grammar~\cite{DBLP:journals/cogsr/SchmidK11}, to name
just a few.  An outstanding challenge, however, is to engineer
program-learners that display the same level of domain-generality as
the humans they are meant to model.

Recent program-learning systems developed within the AI and machine
learning community are impressive along many dimensions, authoring
programs for problem domains like drawing
pictures~\cite{spiral,ellis2017learning}, transforming
text~\cite{gulwani2011automating} and numerical
sequences~\cite{balog2016deepcoder}, and reasoning over common sense
knowledge bases~\cite{muggleton2015meta}.  These systems work in
different ways, but typically hinge upon a carefully hand-engineered
Domain Specific Language (DSL).  The DSL restricts the space of
programs to contain the kinds of concepts needed for one specific
domain.  For example, a picture-drawing DSL could include concepts
like circles and spirals, and a DSL for numerical sequences could
include sorting and reversing lists of numbers.  Modern systems also
learn how to efficiently deploy the DSL on new
problems~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}, but --
unlike human learners -- do not discover the underlying system of
concepts needed to navigate the domain.

We contribute a program-induction system that learns the
domain-specific concepts (DSL) while jointly learning how to use those
concepts.  This joint learning problem models two complementary
notions of domain expertise: (1) domain experts have at their disposal
a powerful, yet specialized repertoire of concepts and abstractions
(analogous to the DSL) while also (2) having accurate intuitions about
when and how to use those concepts to solve new problems.

Our system is called DreamCoder because it acquires these two kinds of
expertise through a novel kind of wake/sleep or `dream'
learning~\cite{hinton1995wake}, iterating through a wake cycle --
where it solves problems by writing programs -- and a pair of sleep
cycles: a sleep cycle that grows the DSL by replaying experiences from
waking and consolidating them into new abstractions (modeled as
fragments of code), and a sleep cycle that improves its knowledge of
how to write programs by training a neural network based on replayed
experiences as well as `dreams', or samples, from the DSL.

DreamCoder builds on two lines of prior work.  One line of work
considers the problem of learning new concepts, abstractions, or
`options' from
experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
while the other line of work considers the problem of learning how to
deploy those concepts
efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
goal with DreamCoder is to show that the combination of these ideas is
uniquely powerful, and pushes us toward program-writing systems that,
like human learners, autonomously acquire the expertise needed to
solve a new class of problems.

\section{Introduction: NIPS}

@@

  \end{figure}

\begin{figure}
\centering
\tiny
\tabcolsep=2pt
\vspace{0pt}
\begin{tabular}{cc}
\begin{tabular}{ccc}
\multicolumn{3}{c}{Tasks:}\\
\hline
1.14531&F&110.9 \\ 
?&CL&163.2 \\
1.29857&F&207.3\\
?&PCFL&143.3 \\
\multicolumn{3}{c}{Learned generative models:}\\
\hline
 \scode{\textbackslash?\textbar(1\textbackslash .\textbackslash d+)}& \scode{((\textbackslash u\textbackslash u)*)\textbar F}& \scode{\textbackslash d\textbackslash d\textbackslash d\textbackslash.\textbackslash d} \\
 \multicolumn{3}{c}{Samples from synthesized generative models:}\\
\hline
1.61&DQDF& 343.8 \\ 
?&F&241.2\\
?&F&647.5\\
1.2&KI&246.8\\
1.5080987&F&728.6\\
?&GL&029.3\\
1.453&F&289.1\\
    \hline
        \\
\end{tabular} 
&
\begin{tabular}{c}
\begin{tabular}{l}
\multicolumn{1}{c}{Learned DSL:}\\
\hline
$f_1() = $ \scode{\textbackslash u\textbackslash w*}\\
$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash u\textbackslash w*)*}\\
$f_3(x) =  f_2(\scode{space}) = $\scode{( \textbar\textbackslash u\textbackslash w*)*}\\
\hline
$f_4(x) = $ \scode{($x$*$x$)}\\
\emph{ (equivalent to regex `plus')}\\
$f_5() = f_4(\scode{\textbackslash l}) = \scode{\textbackslash l*\textbackslash l}$\\
    \hline
\end{tabular}
\\
\\
\includegraphics[width=.4\linewidth]{figures/nampiregexjoshcurve3.eps} 
\end{tabular}
\end{tabular}
\vspace{-.3cm}
\caption{regex stuff}
\end{figure}
